Continuous Deductive Reasoning over Streamed RDF Data

### 2015-11-10 article.tex
gathering of all justifications in example of TA Datalog
expanded discussion of streams and windows in reference to those of the RSP community, and provided definitions of sliding windows and window states

The entailment of the program $\Pi^T$ will be:
\begin{equation}
  cons^T(\Pi^T) = \{\langle c, 1 \rangle, \langle c, 2 \rangle\}
\end{equation}
with the following justifications:
\begin{equation}
  \{ \ \{ \langle a, 1 \rangle, \langle b, 2 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 1 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 2 \rangle \} \  \}
\end{equation}

@@

Having defined the notion of time annotation in Datalog, we now define
the notions of streams and windows before extending to Continuous Datalog.

@@

Streams may be (but are not necessarily) ordered by the instants in
their fact instances, with the instants non-decreasing in the stream,
such that for any two fact instances $\langle f_i, t_i \rangle$ and
$\langle f_j, t_j\rangle$ where $\langle f_i, t_i \rangle \prec
\langle f_j, t_j \rangle$, $t_i < t_j$; we refer to such
streams as {\em in-order streams}, and to streams whose fact instances
are not ordered by their instants as {\em out-of-order streams}.
\end{definition}

\begin{definition}[Window]
\labeldef{continuous datalog: window}

A window of range ${r \in \mathbb{N}}$ over a stream $S$ at an instant
${i \in \mathbb{T}}$ is the \emph{set} of fact instances that appear in
$S$ in the interval ${(i-r,i]}$, written ${W^{S,r}_{i}}$, and defined as follows:
\[
W^{S,r}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in S \land (i - r) < S(\langle f, t\rangle) \leqslant i \}
\]  
\end{definition}

>>>>

The entailment of the program $\Pi^T$ will be:
\begin{equation}
  cons^T(\Pi^T) = \{\langle c, 1 \rangle, \langle c, 2 \rangle\}
\end{equation}
with the following justifications for $\langle c, 1 \rangle$ and $\langle c, 2 \rangle$ respectively:
\begin{gather}
  \{ \ \{ \langle a, 1 \rangle, \langle b, 1 \rangle \} \ \} \\
  \{ \ \{ \langle a, 1 \rangle, \langle b, 2 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 1 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 2 \rangle \} \  \}
\end{gather}

@@

Having defined the notion of time annotation in Datalog, we now define
a notion of streams that is closely related to that currently being
refined by the RSP community.  It differs in two key ways.  Firstly,
elements in streams are not timestamped \emph{RDF graphs} but instances of
individual \emph{RDF triples} (facts). Secondly, fact instances are
annotated for the purposes of distinguishing them from instances of the
same fact asserted/entailed at different times, rather than for the
purpose of ordering them as is the case with timestamped graphs.  This
ordering is instead expressed in the streams themselves, as detailed
formally in \refdef{continuous datalog: stream}, where a fact instance
may occur on many streams and be mapped by each stream to a different
instant.
Each stream in WC-Datalog may be assigned a distinct semantics, in a
similar manner to the distinct predication of timestamps proposed by
the RSP community; they are, therefore, comparably expressive in that
regard, but also represent the manner in which data occurs \emph{in a
stream at a given time} in real-world scenarios.

@@

Streams may be (but are not necessarily) ordered by the instants in
their fact instances, with the instants non-decreasing in the stream,
such that for any two fact instances $\langle f_i, t_i \rangle$ and
$\langle f_j, t_j\rangle$ where $\langle f_i, t_i \rangle \prec
\langle f_j, t_j \rangle$, $t_i < t_j$; we refer to such
streams as {\em in-order streams}, and to streams whose fact instances
are not ordered by their instants as {\em out-of-order streams}.

Based on this definition of streams, we go on to define the notion of
sliding windows over streams, as well as states of those windows at
specific instants.  We do not consider count-based window functions, as
these serve a semantic purpose at high specificity, e.g. \emph{retain
the last ten buses to stop at a specific stop}.  In contrast, as will be
formally described in \refdef{continuous datalog: CDP}, we consider only
the application of the \emph{entirety of a rule set} within the
consecutive states of one (or more) window(s); in this context, an
example application of a count-based window might be to \emph{determine
the implications of the last hundred facts reported by a sensor network,
given some background knowledge regarding that network}.

\begin{definition}[Sliding Window]
\labeldef{continuous datalog: sliding window}

A \emph{sliding-window} definition consists of a stream $S$ that the window is
said to ``slide over'', and a range ${r \in \mathbb{N}}$ that is the
number of consecutive instants that the window covers.  This is written
$W^{S,r}$.
\end{definition}

\begin{definition}[Window State]
\labeldef{continuous datalog: window}

The \emph{state} of a sliding-window $W^{S,r}$ at an instant
${i \in \mathbb{T}}$ is the \emph{set} of fact instances that are
associated in $S$ with any instant in the interval ${(i-r,i]}$, and is
denoted ${W^{S,r}_{i}}$.
The state of a sliding-window is defined as follows:
\[
W^{S,r}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in S \land (i - r) < S(\langle f, t\rangle) \leqslant i \}
\]  
\end{definition}
